---
title: 'STAT232 HW4 '
author: 
- name: Ankit Malhotra
  email: amalh017@ucr.edu
output:
  html_document:
    toc: yes
    toc_depth: 4
---

## **Homework #4 instructions**

**Review textbook [Chapter 5 "Data transformation"](https://r4ds.had.co.nz/transform.html) and [Chapter 7 "Exploratory Data Analysis"](https://r4ds.had.co.nz/exploratory-data-analysis.html), as well as lecture notes on `dplyr` and EDA before answering the homework questions**.

This homework contains 1 question with multiple parts, 100 points in total.

<font color="blue">Replace **INSERT_YOUR_ANSWER** with your own answers.</font>

- First open this `rmd` file in RStudio and click `knnit -> Knit HTML` to render it to HTML format. 
- It is best to read this `rmd` file and the rendered `html` file side-by-side, while you are working on this homework.
- If the question asks you to write some R code, remember to put your code into a **R code chunk**. Make sure both your R code chunk and the output messages / figures are visible in the rendered `html` file.
- For this homework, use **ggplot2** and **dplyr** to visualize and transform your data.

- <font color="blue">**Please comment your R code thoroughly, and follow the R coding style guideline (https://google.github.io/styleguide/Rguide.xml). Partial credit will be deducted for insufficient commenting or poor coding styles.**</font>

- If you have any question about this homework assignment, we encourage you to post it on [**Piazza**].

### **Homework submission guideline**

- Late submission penalties.  
    - Submissions up to 24 hours late will incur a 20% deduction.  
    - Submissions up to 48 hours late will incur a 50% deduction.
    - Submissions more than 48 hours late will receive 0.

- <font color="blue">**If you are using one or both of your free late days, please state here: **</font> **INSERT_YOUR_ANSWER**

- After you complete all questions, save your `rmd` file to `FirstnameLastname-SID-HW4.rmd` and save the rendered `html` or `pdf` file to `FirstnameLastname-SID-HW4.html` or `FirstnameLastname-SID-HW4.pdf`.
- Submit **BOTH your source `rmd` file and the knitted `html` or `pdf` file** to **GradeScope**. Do NOT create a zip file. You can submit multiple times, you last submission will be graded.
***

## **Acknowledgements**

<font color="blue">Please list all the help you have received for completing this homework.</font>

**INSERT_YOUR_ANSWER**

***

**Load necessary packages**

\footnotesize
```{r, collapse=T}
# install.packages("tidyverse")
library(tidyverse) # for `ggplot2` and `dplyr`

# You need to install the nycflights13 package first, then you can comment out the following line.
# install.packages("nycflights13")
library(nycflights13)
```

***

## **Question 1 [100pt] Exploratory data analysis of the `nycflights13` data sets**

In this question, we will take a close look at the `nycflights13::flights` data set which contains all 336,776 flights that departed from New York City in 2013. The data comes from the [US Bureau of Transportation Statistics](http://www.transtats.bts.gov/DatabaseInfo.asp?DB_ID=120&Link=0).

```{r, collapse=T}
?flights # full documentation
# View(flights) # see the data in RStudio Viewer
flights = nycflights13::flights
```

There are some missing values in the data set which were caused by **canceled flights**. You can clean the `flights` data by removing flight record that has missing values in `dep_delay` or `arr_delay`, and save the **non-canceled** flights in a new tibble `not_canceled`.

```{r, collapse=T}
not_canceled <- filter(flights, !is.na(dep_delay), !is.na(arr_delay))
not_canceled
```  

In the following questions, you will use `ggplot2` and `dplyr` functions to perform exploratory data analysis. 

<font color="blue">If not specified, flight delays usually refer to **arrival delays**</font>.

### (a) [15pt] Pick observations by their values - `filter()`

Find the **non-canceled** flights that satisfy each of the following conditions.

(i) Had an arrival delay of two or more hours

```{r}
filter(not_canceled, arr_delay >= 120)
```


(ii) Were operated by United, American, or Delta

```{r}
filter(not_canceled, carrier %in% c("AA", "DL", "UA"))
```

(iii) Arrived more than two hours late, but didn't leave late

```{r}
filter(not_canceled, arr_delay > 120 & dep_delay <= 0)
```


(iv) Departed between midnight and 6am (inclusive)

```{r}
filter(not_canceled, dep_time %in% c(0000:0600) | dep_time == 2400)
```


***

### (b) [10pt] Reorder the rows - `arrange()`

Find the following **non-canceled** flights. 

(i) Find the top five most delayed flights. Report the flight date, origin, destination, carrier, flight number, and  arrival delay.

```{r}
head(select(arrange(not_canceled, desc(arr_delay)), year, month, day, origin, dest, carrier, flight, arr_delay), 5)
```


(ii) Find the flights traveled the longest by distance. What are the carrier, the origin airport, and the destination airport of these flights?

```{r}
select(arrange(not_canceled, desc(distance)), flight, carrier, origin, dest)
```


***

### (c) [5pt] Pick variables by their names - `select()`

List at least three ways to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from **non-canceled** flights.

```{r}
#1st way :
col_names <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
non_cancelled_subset_1 <- select(not_canceled, one_of(col_names))
non_cancelled_subset_1

#2nd way :
select(not_canceled, "dep_time", "dep_delay", "arr_time", "arr_delay")

#3rd Way : 
non_canceled_subset_2 <- select(not_canceled, starts_with("dep"), starts_with("arr"))
non_canceled_subset_2
```


***

### (d) [20pt] Create new variables with functions of existing variables - `mutate()`

(i) For the **non-canceled** flights, use the `mutate()` function to create a new variable `flight_time` which equals `arr_time` - `dep_time`. Make a scatter plot to compare `flight_time` with `air_time`.

```{r}
library(ggplot2)
not_canceled <- mutate(not_canceled, flight_time = arr_time - dep_time)

not_canceled

ggplot(data = not_canceled, mapping = aes(x = flight_time, y = arr_time)) +
  geom_point()
```



(ii) What do you expect to see? What do you see? Why are there negative values in `flight_time`?

I expected to see a scatter plot which shows some kind of uniform simulation of points. I see different clusters of scatters which is odd. 
(iii) Fix your R code to correct the negative values in `flight_time`. Make the scatter plot again. 

**Hint:** The `lubridate` package would be helpful. Check out our textbook **[Chapter 16 "Dates and times"](http://r4ds.had.co.nz/dates-and-times.html)**

```{r}
library(lubridate)

# Convert dep_time and arr_time to POSIXct format
not_canceled$dep_time_datetime <- as.POSIXct(paste(not_canceled$year, not_canceled$month, not_canceled$day, not_canceled$dep_time, sep = "-"), format = "%Y-%m-%d-%H%M")
not_canceled$arr_time_datetime <- as.POSIXct(paste(not_canceled$year, not_canceled$month, not_canceled$day, not_canceled$arr_time, sep = "-"), format = "%Y-%m-%d-%H%M")

# Calculate flight time
not_canceled$flight_time <- as.numeric(difftime(not_canceled$arr_time_datetime, not_canceled$dep_time_datetime, units = "mins"))

# Remove negative flight times
not_canceled <- not_canceled %>% filter(flight_time >= 0)

ggplot(data = not_canceled, aes(x = air_time, y = flight_time)) +
  geom_point() +
  labs(x = "Distance (miles)", y = "Flight time (minutes)", title = "Flight time vs. Distance")

not_canceled

```


***

### (e) **[Extra Credit 20pt]** More on date and time

You have probably wondered why your scatter plot didn't have a clear linear relationship. The reasons were due to the time zone differences as well as the daylight saving time (DST). Modify your R code to solve these issues. 

**Hint:** Check out the `nycflights13::airports` data set. You can _join_ `flights` and `airports` using destination airport FAA code as the _key_.

```{r, collapse=T}
#?airports
airports %>% glimpse()

left_join(flights, airports, by=c("dest"="faa")) %>% glimpse()
```

**INSERT_YOUR_ANSWER**

```{r}
flights_edt <- flights %>%
  mutate(
    dep_time_datetime = with_tz(as.POSIXct(paste(year, month, day, dep_time, sep = "-"), format = "%Y-%m-%d-%H%M"), tzone = "EST5EDT"),
    arr_time_datetime = with_tz(as.POSIXct(paste(year, month, day, arr_time, sep = "-"), format = "%Y-%m-%d-%H%M"), tzone = "EST5EDT")
  )

# Join flights and airports using destination airport FAA code as the key
flights_airports <- flights_edt %>%
  left_join(airports, by = c("dest" = "faa"))

# Calculate flight time
# flights_airports$flight_time <- as.numeric(difftime(flights_airports$arr_time_datetime, flights_airports$dep_time_datetime, units = "mins"))

flights_airports <- flights_airports %>%
  mutate(sched_dep_time = dep_time_datetime + as.numeric(tz))

# Taking into account daylight saving time
flights_airports <- flights_airports %>%
  mutate(sched_dep_time = sched_dep_time - hours(ifelse(dst == "A", 1, 0)))

flights_airports <- flights_airports %>%
  mutate(sched_arr_time = arr_time_datetime + as.numeric(tz))

# Taking into account daylight saving time
flights_airports <- flights_airports %>%
  mutate(sched_arr_time = sched_arr_time - hours(ifelse(dst == "A", 1, 0)))

flights_airports <- mutate(flights_airports, flight_time = arr_time - dep_time)

# Scatter plot of flight time vs. distance
ggplot(data = flights_airports, aes(x = distance, y = flight_time)) +
  geom_point() +
  labs(x = "Distance (miles)", y = "Flight time (minutes)", title = "Flight time vs. Distance")
```



***

### (f) [50pt] Calculate group-wise summaries - combine `summarise()` and `group_by()`

#### (f.1) [25pt] Daily flight cancellation rate

(i) Suppose we are interested in studying the number of canceled flights per day. Is there a pattern? 
Make a scatter plot of the proportion of canceled flights per day (x-axis) vs average delay per day (y-axis). 
Use point size to represent the number of originally scheduled flights on each day. 
Add a smoothed fitted line to your scatter plot. 
Is the proportion of canceled flights related to the average delay?

**Hint:** Group `flights` (do not use `not_canceled`) by both month and day, so that your scatter plot has 365 points in total.

**INSERT_YOUR_ANSWER**
```{r}
# group_by(flights, month, day) %>% summarise(count = n()) %>% arrange(desc(count))
library(dplyr)



# USE THE METHOD FOR FINDING CANCELLED FLIGHTS ABOVE
# cancelled_flights <- filter(flights, is.na(dep_delay) | is.na(arr_delay)) 

f1 <- mutate(flights,
             canceled = ifelse(is.na(dep_delay) | is.na(arr_delay), 1, 0)) %>%
      group_by(month,day) %>%
      summarise(daily_canceled = sum(canceled) / n(), delay = mean(arr_delay, na.rm = T), scheduled = n(), .groups = 'drop')

ggplot(data=f1) +
  geom_point(mapping=aes(x=daily_canceled, y=delay, size=scheduled), alpha=1/3) +
  geom_smooth(mapping=aes(x=daily_canceled, y=delay), se=FALSE) +
  xlab("proportion of canceled flights per day") +
  ylab("average delay per day")

f2 <- arrange(f1, desc(daily_canceled))
f2 <- f2[3:nrow(f2),]

ggplot(data=f2) +
  geom_point(mapping=aes(x=daily_canceled, y=delay, size=scheduled), alpha=1/3) +
  geom_smooth(mapping=aes(x=daily_canceled, y=delay), se=FALSE) +
  xlab("proportion of canceled flights per day") +
  ylab("average delay per day")


# Group flights by month and day
# flights_daily <- flights %>%
#   group_by(year, month, day) %>%
#   summarise(
#     canceled = sum(is.na(dep_time), na.rm = TRUE),
#     scheduled = n(),
#     avg_delay = mean(dep_delay, na.rm = TRUE)
#   ) %>%
#   ungroup() %>%
#   mutate(
#     canceled_prop = canceled / scheduled,
#     avg_delay = ifelse(scheduled == 0, NA, avg_delay),
#     canceled_prop = ifelse(scheduled == 0, NA, canceled_prop)
#   )
# 
# # Scatter plot of cancellation rate vs. average delay
# ggplot(data = flights_daily, aes(x = canceled_prop, y = avg_delay)) +
#   geom_point(aes(size = scheduled)) +
#   geom_smooth(method = "loess") +
#   labs(x = "Proportion of canceled flights", y = "Average delay (minutes)", size = "Number of scheduled flights", title = "Daily flight cancellation rate vs. average delay")
```


(ii) Which two days had the highest flight cancellation rates?

13th May and 21st of august had the most cancellation rates
(iii) Remove the two days with the highest flight cancellation rates and re-draw your scatter plot.

**INSERT_YOUR_ANSWER**

(iv) What do you think might be the main reason for these two high cancellation rates? Find out supporting evidences from the `nycfights13::weather` data set.

```{r, collapse=T}
#?weather
weather %>% glimpse()
```

**INSERT_YOUR_ANSWER**

***

#### (f.2) [25pt] Best/worst carrier of the year

(i) Suppose we are interested in identifying the carriers with the least/worst delays. To disentangle the effects of bad airports vs. bad carriers, we group all **non-canceled** flights by both carrier and dest, then calculate the average arrival delay for each carrier and dest pair. Make side-by-side boxplots to compare the distribution of average arrival delay among different carriers. Flip your boxplot to have horizontal display and rank all boxes by their median values.

```{r}
grouping_data <- group_by(not_canceled, carrier, dest) %>% summarise(avg_arr_delay = mean(arr_delay), count_each = n()) %>% arrange(desc(avg_arr_delay))

ggplot(data = grouping_data, mapping = aes(x = reorder(carrier, avg_arr_delay, FUN = median), y = avg_arr_delay, fill = carrier)) + 
  geom_boxplot()
```


(ii) Which carrier had the least delays? Which carrier had the worst delays? 

=> "AS" had the least delays, whereas, carrier "UA" had the worst delays, given that it has the most number of outliers. 

(iii) Is there a small `n` problem in your data analysis? Count the number of destinations per carrier.

**Hint:** Check out the `count()` function @ https://dplyr.tidyverse.org/reference/tally.html

Yes there is a problem, where the number of aggregated records are a mismatch.
This can probably be because of the number of destinations for a few records being only 1.



```{r}
flights_new <- not_canceled %>%
  group_by(carrier) %>%
  summarize(diff_dest = n_distinct(dest))

flights_new

```


(iv) Remove the carriers that only flew to one destination. Re-draw your boxplot and re-name the best/worst carriers.

```{r}
grouping_data_after_1 <- group_by(not_canceled, carrier, dest) %>% filter(!(carrier %in% c("AS", "HA", "F9"))) %>% summarise(avg_arr_delay = mean(arr_delay), count_each = n()) %>% arrange(desc(avg_arr_delay))

ggplot(data = grouping_data_after_1, mapping = aes(x = reorder(carrier, avg_arr_delay, FUN = median), y = avg_arr_delay, fill = carrier)) + 
  geom_boxplot()

```
1. Least Delays (acc to plots) : -

  DL - Carrier

2. Worst Delays : - 

  FL - Carrier
