---
title: 'STAT232 HW5'
author: 
- name: Ankit Malhotra
  email: amalh017@ucr.edu
output:
  html_document:
    toc: yes
    toc_depth: 4
---

## **Homework #5 instructions**

**Review [the ISLR textbook](https://www.statlearning.com) Chapter 3 "Linear Regression", as well as lecture notes on map visualization and linear regression before answering the homework questions**.

This homework contains 2 questions, each with multiple parts, 100 points in total.  

<font color="blue">Replace **INSERT_YOUR_ANSWER** with your own answers.</font>

- First open this `rmd` file in RStudio and click `Knit -> Knit to HTML` to render it to HTML format. 
- It is best to read this `rmd` file and the rendered `html` file side-by-side, while you are working on this homework.
- If the question asks you to write some R code, remember to put your code into a **R code chunk**. Make sure both your R code chunk and the output messages / figures are visible in the rendered `html` file.
- For this homework, use **ggplot2** to visualize your data.

- <font color="blue">**Please comment your R code thoroughly, and follow the R coding style guideline (https://google.github.io/styleguide/Rguide.xml). Partial credit will be deducted for insufficient commenting or poor coding styles.**</font>

- If you have any question about this homework assignment, we encourage you to post it on [**Piazza**].

### **Homework submission guideline**

- Late submission penalties.  
    - Submissions up to 24 hours late will incur a 20% deduction.  
    - Submissions up to 48 hours late will incur a 50% deduction.
    - Submissions more than 48 hours late will receive 0.

- <font color="blue">**If you are using one or both of your free late days, please state here: **</font> **INSERT_YOUR_ANSWER**

- After you complete all questions, save your `rmd` file to `FirstnameLastname-SID-HW5.rmd` and save the rendered `html` or `pdf` file to `FirstnameLastname-SID-HW5.html` or `FirstnameLastname-SID-HW5.pdf`.
- Submit **BOTH your source `rmd` file and the knitted `html` or `pdf` file** to **GradeScope**. Do NOT create a zip file. You can submit multiple times, you last submission will be graded.

***

## **Acknowledgements**

<font color="blue">Please list all the help you have received for completing this homework.</font>

**INSERT_YOUR_ANSWER**

***

**Install necessary packages**

Note that you only need to install each package once. Then you can comment out the following installation lines.

``````{r}
#install.packages("maps")d`
#install.packages("mapproj")
```

***

**Load necessary packages**

```{r, collapse=T}
library(tidyverse) # for `ggplot2`, `dplyr`, and more

library(maps) # for map visualization
#library(mapdata)
library(mapproj) 

library(datasets) # for `state` data set
library(nycflights13) # for the 2013 NYC flights data set
```

***

## **Question 1 [45pt] Map visualization with `ggplot2`**

Recall in HW1, we used the `maps` package to generate map visualizations. In this homework, we will draw maps with `ggplot2`. 

The `maps` package comes with a plotting function, but, we will opt to use `ggplot2` functions (`geom_polygon` and `geom_map`) to plot the maps in the `maps` package.

Recall that `ggplot2` operates on data frames. We will use the `map_data()` function (provided by `ggplot2`), which turns a series of points along an outline into a data frame of those points.

### (a) [5pt] Plot the USA map

First we load the USA map from `maps`.
```{r, collapse=T}
usa_map <- map_data("usa")
dim(usa_map)
glimpse(usa_map)
```

In the lecture, we have learned how to use `geom_polygon()` to make a simple black map (no line color, but with a black fill).

```{r}
ggplot(data = usa_map) + 
  geom_polygon(aes(x = long, y = lat, group = group)) + 
  coord_quickmap()
```

Google the coordinates of UCR, and then use `geom_point()` to mark the location of UCR on the USA map.
In addition, change the outline/border color as well as the fill-in color of your map.

**INSERT_YOUR_ANSWER**
```{r}
ggplot(data = usa_map) + 
  geom_polygon(aes(x = long, y = lat, group = group), color = "blue", fill = "lightgray") + 
  geom_point(x = -117.3424, y = 33.9556, color = "red", size = 3) +
  coord_quickmap()
```


***

### (b) [15pt] Plot the states map

In addition to `geom_polygon()`, we can use `geom_map()` to draw maps too. Basically, `geom_map()` acts as a wrapper of `geom_polygon()`. See more details in the `geom_map()` documentation @ http://ggplot2.tidyverse.org/reference/geom_map.html

Here is the example code of a states map from the lecture. 
We can plot all the states, map the `fill` aesthetic to `region` and set the the lines of state borders to white color.

```{r, collapse=T}
states_map <- map_data("state")
dim(states_map)
glimpse(states_map)
```

```{r, collapse=T}
ggplot() + 
  geom_map(data = states_map, map = states_map,
           aes(map_id = region, fill = region), color="white") + 
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_quickmap() +
  guides(fill = FALSE) # do this to leave off the color legend
```

Next, we will use the built-in `state` data sets in R to annotate our states map. In particular, `state.x77` is a two-dimensional array containing 8 variables and data from all 50 states.

```{r, collapse=T}
?state.x77
head(state.x77)

state_data <- as.data.frame(state.x77)
state_data$State <- tolower(rownames(state_data))
state_data %>% glimpse()
```

Below is the example code from the lecture for a state population map. We first create an aesthetic mapping for `map_id` to the column `State` (state names in lower case) in the `state_data` data frame. We then call `geom_map` again and map the `fill` aesthetic to the `Population` variable in `state_data`.

```{r, collapse=T}
ggplot(data = state_data, aes(map_id = State))+ 
  geom_map(map = states_map,
           aes(fill = Population)) + 
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_quickmap() 
```

#### (i) [5pt] Draw your own states map and map the `fill` aesthetic to `Income` in the `state.x77` data set

**INSERT_YOUR_ANSWER**
```{r}
ggplot(data = state_data, aes(map_id = State))+ 
  geom_map(map = states_map,
           aes(fill = Income)) + 
  expand_limits(x = states_map$long, y = states_map$lat) +
  coord_quickmap() 
```


***

#### (ii) [10pt] Add 50 colorful points to your map 
Use one point to mark one state (state coordinates can be found in `state.center`). Map the color of the points to `state.region`. Map the `size` aesthetic of the points to `Population`.

```{r, collapse=T}
head(state.center)
head(state.region)
```

**INSERT_YOUR_ANSWER**
```{r}
ggplot(data = state_data, aes(map_id = State)) +
  geom_map(map = states_map,
           aes(fill = Income, color = state.region)) +
  # geom_point(data = state.center,
  #             aes(x = long, y = lat, color = state.region, size = Population)) +
  geom_point(aes(x = state.center$x, y = state.center$y, color = state.region, size = Population))
  expand_limits(x = state.center$x, y = state.center$y)
```


***

### (c) [25pt] NYC flight destination map

Recall `nycflights13::flights` data set contains all 336,776 flights that departed from New York City in 2013.

```{r, collapse=T}
?flights # full documentation
glimpse(flights)
```

#### (i) [10pt] Count the number of flights per destination

How many unique destination airports did these NYC flights connected to? How many **non-canceled** flights per destination? Which destination had the largest number of arrival flights from NYC? Which destination had the smallest number of arrival flights from NYC?

**Note:** Recall in HW4 we have cleaned the `flights` data by removing flight record that has missing values in `dep_delay` or `arr_delay`, and save the result as the **non-canceled** flights.

```{r}
not_canceled <- filter(flights, !is.na(dep_delay), !is.na(arr_delay))

not_canceled <- not_canceled %>% 
  group_by(dest) %>%
  summarize(num_flights = n()) %>% arrange(num_flights)
not_canceled
```

Destination with the largest number of arrival flights from NYC: ATL 
Destination with the smallest number of arrival flights from NYC: LEX
*** 

#### (ii) [15pt] Mark all destination airports on a states map
Find out the coordinates of the destination airports from `nycflights13::airports`. Draw each destination airport as a point on a states map, and map a point aesthetic to the number of **non-canceled** flights flew to that destination from NYC in 2013.

**Hint:** Suppose that you have saved your list of unique destinations and their corresponding flight counts in a tibble called `per_dest`. Then you can _join_ `per_dest` and `airports` using destination airport FAA code as the _key_.

```{r, eval=F}
#?airports
airports %>% glimpse()
# left_join(per_dest, airports, by=c("dest"="faa"))
```

**INSERT_YOUR_ANSWER**

```{r}
dest_map <- not_canceled %>%
  left_join(airports, by = c("dest" = "faa"))

states_map <- map_data("state")

ggplot() +
  geom_polygon(data = states_map, aes(x = long, y = lat, group = group), fill = "gray80") +
  geom_point(data = dest_map, aes(x = lon, y = lat, size = num_flights), color = "dodgerblue", alpha = 0.7) +
  scale_size_continuous(range = c(1, 10), name = "Non-canceled Flights from NYC in 2013") +
  labs(x = "Longitude", y = "Latitude", title = "Destination Airports from NYC in 2013") +
  theme_bw()
```


Pay attention to the warning message. Which destination airports have missing values?
```{r}
missing_dests <- dest_map %>%
  filter(is.na(lat) | is.na(lon)) %>%
  select(dest) %>%
  distinct()
missing_dests
```


**
The below 4 have missing values : 

BQN				
PSE				
SJU				
STT				
**

***

## **Question 2 [55pt] Simple linear regression**

### (a) [10pt] Population regression line vs least squares regression line

In this question, we will reproduce the left panel of Figure 3.3 in the **ISLR** textbook.

Write your own R code the simulate 100 data points from the following model.

$$Y=2+3X+\epsilon$$
$$X \sim Uniform(A=-2, B=2)$$
$$\epsilon \sim N(\mu=0, \sigma=2)$$

Then use `ggplot2` to make a scatter plot of the 100 data points $(x_1,y_1), \cdots (x_{100},y_{100})$. Add the population regression line $Y=2+3X$ in red color; add the least squares fit of your simulated data as a blue line. 

**INSERT_YOUR_ANSWER**
```{r}
set.seed(1997)

X <- runif(100, min = -2, max = 2)

epsilon <- rnorm(100, mean = 0, sd = 2)

Y <- 2 + 3 * X + epsilon

data <- data.frame(X, Y)

# # Plot the data points and the population regression line
# plot(data, main = "Population Regression Line vs Least Squares Regression Line",
#      xlab = "X", ylab = "Y", xlim = c(-2, 2), ylim = c(-6, 10))
# abline(a = 2, b = 3, col = "red", lwd = 2, lty = 2)

scatter_plot <- ggplot(data, aes(x = X, y = Y)) +
  geom_point(alpha = 0.5) +
  xlab("X") +
  ylab("Y") +
  theme_minimal()

scatter_plot <- scatter_plot +
  geom_abline(intercept = 2, slope = 3, color = "red", linetype = "dashed")

scatter_plot <- scatter_plot +
  stat_smooth(method = "lm", color = "blue", se = FALSE)

print(scatter_plot)
```


```{r, collapse=T}
# Always remember to set the random seed for simulation studies
set.seed(1997) # feel free to change 232 to your lucky number
```

***

### (b) [10pt] Estimate the coefficients $\beta_0$ and $\beta_1$ 

In the lecture, we have learned that the analytical solution of the least squares regression $Y = \beta_0 + \beta_1 X + \epsilon$ is

$$ 
\begin{aligned}
 \hat{\beta_1} & = 
 \frac{\sum_{i=1}^{n} (x_i - \bar{x})(y_i - \bar{y})}
 {\sum_{i=1}^{n} (x_i - \bar{x})^2} \\
 \hat{\beta_0} & = \bar{y} - \hat{\beta_1} \bar{x}
\end{aligned} 
$$

Write your own R code to calculate $\hat{\beta_0}$ and $\hat{\beta_1}$ using the above equations. Compare your results with the coefficients calculated by the `lm` function. Are they the same?

**INSERT_YOUR_ANSWER**
```{r}
# Calculate the sample mean of X and Y
x_mean <- mean(data$X)
y_mean <- mean(data$Y)

# Calculate the numerator and denominator of the formula for beta1
num <- sum((data$X - x_mean) * (data$Y - y_mean))
den <- sum((data$X - x_mean)^2)

# Calculate beta1
beta1_hat <- num / den

# Calculate beta0
beta0_hat <- y_mean - beta1_hat * x_mean

# Display the estimated coefficients
cat("Estimated coefficients using the formula:\n")
cat("beta0 =", round(beta0_hat, 3), "\n")
cat("beta1 =", round(beta1_hat, 3), "\n")
```

```{r}
# Calculate the coefficients using the lm function
coefs <- lm(Y ~ X, data = data)$coefficients

# Print the coefficients
print(paste("beta0 (lm) =", coefs[1]))
print(paste("beta1 (lm) =", coefs[2]))
```

Yes they are the same!


***

### (c) [20pt] Assess accuracy of the regression model

Recall that the accuracy of a linear regression fit is typically assessed using the **residual standard error (RSE)** and the **$R^2$ statistic**.

```{r}
# Set seed for reproducibility
set.seed(123)

# Generate predictor variable
x <- runif(100, min = -2, max = 2)

# Generate response variable
y <- 2 + 3 * x + rnorm(100, mean = 0, sd = 2)

# Fit linear regression model
model <- lm(y ~ x)

# Calculate RSE
rse <- summary(model)$sigma
cat("Residual standard error (RSE):", rse, "\n")

# Calculate R-squared
r_squared <- summary(model)$r.squared
cat("R-squared ($R^2$):", r_squared, "\n")
```


#### (i) [10pt] Calculate the residual standard error (RSE)

Residual standard error (RSE) is an estimate of $\sigma$, the standard deviation of $\epsilon$.

$$
\begin{aligned}
\text{RSE} = \hat{\sigma} &= \sqrt{\frac{1}{n-2}\text{RSS}} \\
&= \sqrt{\frac{1}{n-2} \sum_{i=1}^{n} ( y_i - \hat{y_i} )^2} 
\end{aligned}
$$ 
Write your own R code to calculate RSE using the above equation. Compare your result with the residual standard error calculated by the `lm` function. Are they the same? How does the RSE value relate to our noise model $\epsilon \sim N(\mu=0, \sigma=2)$?

**INSERT_YOUR_ANSWER**
```{r}
# Calculate RSS
rss <- sum((y - predict(model))^2)

# Calculate RSE
rse_manual <- sqrt(rss / (length(x) - 2))
cat("Residual standard error (RSE) - manual calculation:", rse_manual, "\n")
```

```{r}
set.seed(123)

x <- runif(100, min = -2, max = 2)

y <- 2 + 3 * x + rnorm(100, mean = 0, sd = 2)

# Fit linear regression model
model <- lm(y ~ x)

# Calculate RSE
rse <- summary(model)$sigma
cat("Residual standard error (RSE):", rse, "\n")
```

Yes, it is the same.

***

#### (ii) [10pt] Calculate the $R^2$ statistics

Write your own R code to compute the $R^2$ value using the following equations.

$$
\begin{aligned}
\text{TSS} & = \sum_{i=1}^{n} (y_i - \bar{y})^2 \\
\text{RSS} & = \sum_{i=1}^{n} \left(y_i - \hat{f}(x_i) \right)^2 \\
R^2 & = \frac{\text{TSS}-\text{RSS}}{\text{TSS}} = 1 - \frac{\text{RSS}}{\text{TSS}} \\
\end{aligned}
$$
Compare your result with the R-squared value calculated by the `lm` function. Are they the same? 

**INSERT_YOUR_ANSWER**

```{r}
# Calculate TSS
tss <- sum((y - mean(y))^2)

# Calculate RSS
rss <- sum((y - predict(model))^2)

# Calculate R-squared
r_squared <- 1 - rss / tss
cat("R-squared ($R^2$):", r_squared, "\n")
```

```{r}
# Calculate R-squared using lm() function
cat("R-squared ($R^2$) using lm():", summary(model)$r.squared, "\n")
```

Yes, it is the same.

***

### (d) [15pt] Simulations of least squares regression

Now, let's repeat the simulation 10 times, and reproduce the right panel of Figure 3.3 in the **ISLR** textbook.

**Notes**: 

- You can draw all the least squares lines using the same blue color.
- Write your code using a `for` loop, rather than manually copying and pasting your code 10 times.
- You can save your ggplot to a variable (e.g. `gg`) and keep adding new layers inside your for loop. So the pseudocode will look like:

```{r, eval=F}
gg <- ggplot() +
  geom_abline(draw the population regression line)

for ( every simulation ) {
  ## add your simulation code here
  gg <- gg + 
    geom_smooth(draw the sample regression line) ## alternatively you can use geom_abline()
}

gg ## call gg at the end to plot it.
```

**INSERT_YOUR_ANSWER**

```{r}
population_line <- lm(y ~ x)

gg <- ggplot() +
  geom_abline(intercept = coef(population_line)[1], slope = coef(population_line)[2], color = "red", linetype = "dashed") +
  xlim(-2, 2) +
  ylim(-6, 10) +
  xlab("x") +
  ylab("y")


for (i in 1:10) {
  y <- 2 + 3 * x + rnorm(100, mean = 0, sd = 2)
  sample_line <- lm(y ~ x)
  gg <- gg +
    geom_abline(intercept = coef(sample_line)[1], slope = coef(sample_line)[2], color = "blue")
}

# Plot ggplot
print(gg)
```


***