---
title: "Intro to GGplot2"
output: html_notebook
---

FOR ADDING A PICTURE, ADD THE PICTURE TO THE LOCAL FOLDER AND ADD THE PATH / URL OF THE PICTURE TO R AND PUT IT IN PARENTHESIS AND WE ARE GOOD


```{r}
library(ggplot2)
library(tidyverse)

?mpg

head(mpg)

glimpse(mpg)

ggplot(data = mpg) +  # creates a canvas / adds just a skeleton to the actual graph we eentually want to build
  geom_point(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) +  # creates a canvas / adds just a skeleton to the actual graph we eentually want to build
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) # class is a column / data point / variable in the dataset, and when we say
# color = class, it takes all the unique values of that column and colors it differently

ggplot(data = mpg) +  # creates a canvas / adds just a skeleton to the actual graph we eentually want to build
  geom_point(mapping = aes(x = displ, y = hwy, size = class)) # class is a column / data point / variable in the dataset, and when we say
# size = class, it takes all the unique values of that column and sizes it differently

ggplot(data = mpg) +  # creates a canvas / adds just a skeleton to the actual graph we eentually want to build
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class)) # class is a column / data point / variable in the dataset, and when we say
# alpha = class, it takes all the unique values of that column and uses different levels of transparency with it.

ggplot(data = mpg) +  # creates a canvas / adds just a skeleton to the actual graph we eentually want to build
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class, color = class)) # class is a column / data point / variable in the dataset, and when we say
# alpha = class, it takes all the unique values of that column and uses different levels of transparency with it.


ggplot(data = mpg) +  # creates a canvas / adds just a skeleton to the actual graph we eentually want to build
  geom_point(mapping = aes(x = displ, y = hwy, shape = class)) # class is a column / data point / variable in the dataset, and when we say
# shape = class, it takes all the unique values of that column and uses different shapes built in with ggplot2.

# only 6 DIFFERENT shapes available in R

ggplot(data = mpg) +  # creates a canvas / adds just a skeleton to the actual graph we eentually want to build
  geom_point(mapping = aes(x = displ, y = hwy)) +
             facet_wrap(~ class, nrow = 2)
#facet_wrap creates different graphs for all the unique values in variables
# nrow basically puts the different graphs in  the number of rows mentioned in the code above

ggplot(data = mpg) +  # creates a canvas / adds just a skeleton to the actual graph we eentually want to build
  geom_point(mapping = aes(x = displ, y = hwy)) +
             facet_grid(drv ~ class)
# basically creates subgrpoups of data based on the parameter
```
```{r}
library(gridExtra)

plot1 <- ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
plot2 <- ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))

grid.arrange(plot1, plot2, ncol=2)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, col = drv)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, col = drv))



ggplot(data = mpg, mapping = aes(x=displ, y=hwy, col = drv)) +
  geom_point() + 
  geom_smooth(mapping = aes(linetype = drv))

ggplot(data = mpg, mapping = aes(x=displ, y=hwy, col = drv)) +
  geom_point() + 
  geom_smooth(se=FALSE)

ggplot(data = mpg, mapping = aes(x=displ, y=hwy, col = drv)) +
  geom_point() + 
  geom_smooth(method = "lm")


```

trying ggplot for bars

```{r}

head(diamonds)

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```


```{r}
library(gridExtra)
gg <- ggplot(data = diamonds, mapping = aes(x=cut))
plot1 <- gg + geom_bar()
plot2 <- gg + stat_count()
grid.arrange(plot1, plot2, ncol = 2)
```

```{r}
demo <- tribble(
  ~a, ~b,
  "bar_1", 20,
  "bar_2", 30,
  "bar_3", 40
)
gg <- ggplot(data = demo, mapping = aes(x=a))
plot1 <- gg + geom_bar()
plot2 <- gg + geom_bar(mapping = aes(y = b), stat = "identity")
plot3 <- gg + geom_col(mapping = aes(y = b))
grid.arrange(plot1, plot2, plot3, ncol = 3)
```
```{r}
plot11 <- ggplot(data = diamonds) + geom_bar(mapping = aes(x=cut, y = after_stat(prop), group = 1, fill = cut))
plot12 <- ggplot(data = diamonds) + geom_bar(mapping = aes(x=cut, y = after_stat(prop)))
plot13 <- ggplot(data = diamonds) + geom_bar(mapping = aes(x=cut, fill = cut))
plot14 <- ggplot(data = diamonds) + geom_bar(mapping = aes(x=cut, color = cut))
grid.arrange(plot11, plot12, plot13, plot14, ncol=2)
```

```{r}
plot20 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, color = clarity, fill = clarity))

plot21 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, color = clarity, fill = clarity), position = "identity") # not the ideal way to draw a graph


grid.arrange(plot20, plot21, ncol = 2)
```

```{r}
plot30 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, color = clarity, fill = clarity), alpha = 1/5, position = "identity")

plot31 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, color = clarity, fill = clarity), position = "identity", fill = NA) # not the ideal way to draw a graph either

plot32 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")


grid.arrange(plot30, plot31, plot32, ncol = 1)
```

```{r}
plot33 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
plot33
```

```{r}
gg <- ggplot(data = diamonds, mapping = aes(x=carat)) +
  theme(legend.position = "bottom", legend.text = element_text(size = 5))
plot35 <- gg + geom_histogram(aes(fill=cut), binwidth=0.5)
plot36 <- gg + geom_freqpoly(aes(color=cut), binwidth=0.5)
grid.arrange(plot35, plot36, ncol=2)
```

```{r}
ggplot(data = diamonds) + 
  geom_histogram(aes(x=carat, fill=cut), binwidth = 0.5, position = "dodge") +
  theme(legend.position = "bottom")
```

```{r}
ggplot(data = diamonds) + 
  geom_freqpoly(aes(x=carat, y=after_stat(density), color=cut), binwidth = 0.5) +
  theme(legend.position = "bottom")
```

```{r}
ggplot(data = diamonds, mapping = aes(x=carat)) + 
  geom_histogram(aes(y=after_stat(density)), binwidth = 0.1, color = "black", fill = "white") +
  geom_density(alpha = 0.2, fill = "#FF6666")
```
**BOXPLOT**

```{r}
ggplot(data=mpg) + 
  geom_boxplot(mapping = aes(x = year, y = hwy))
```

```{r}
ggplot(data=mpg) + 
  geom_boxplot(mapping = aes(x = class, y = hwy))
```
```{r}
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x=reorder(class, hwy, FUN = median), y = hwy)) + 
  xlab('class')
  ylab('highway fuel efficiency')
```

```{r}
ggplot(data = mpg) + 
  geom_boxplot(mapping = aes(x=reorder(class, hwy, FUN = median), y = hwy, fill = class)) + 
  xlab('class') +
  coord_flip() + 
  legend.position = None
```

```{r}
ggplot(data = mpg) + 
  geom_violin(mapping = aes(x = class, y = hwy))
```
## INSTALL PACKAGES FOR DIFFERENT COLOR PALLETTES TO MAKE IT LOOK MORE VISUALLY APpealing


```{r}

ggplot(data = count(x = diamonds, cut, clarity)) + 
  geom_tile(mapping = aes(x=cut, y=clarity, fill = n))






```

What is n? Is it the default column name given to aggregated outputs?

Can we make like an environment or a file which renders all the packages we need at once? like requirements.txt in python    

So using jitter doesnt alter the accuracy of data?

```{r}
library(gridExtra)

plot101 <- ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
plot102 <- ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
grid.arrange(plot101, plot102, ncol = 2)
```



```{r}
set.seed(123)
plot201 <- ggplot(data = mpg, mapping = aes(x = drv, y = hwy)) + 
  geom_point(position = "jitter")
plot202 <- ggplot(data = mpg, mapping = aes(x = drv, y = hwy)) + 
  geom_jitter()
grid.arrange(plot201, plot202, ncol = 2)
```


Controlling the jittering

```{r}
set.seed(999)
p1 <- ggplot(data = mpg, mapping = aes(x=drv, y=displ)) +
  geom_point()
p2 <- p1 + geom_jitter(col="red")
p3 <- p1 + geom_jitter(col = "green", width = 0.2, height = 4)
grid.arrange(p2, p3, ncol = 2)
```




